Parsing Postfix log files

### 2008-05-20 doc/logparser-background.tex
:%s/logs/log files/gc

This section provides background information helpful in understanding the
remainder of the paper.  It begins with a discussion of the motivation
underlying the project, followed by some technical information: the use of
a database as an \API{}\@; a brief introduction to \SMTP{}\@; a longer
introduction to Postfix, concentrating on the topics most relevant to this
paper, namely Postfix anti-spam restrictions and policy servers.  The
assumptions made in designing and implementing the parser are explained, as
are the conventions used in this paper.  Other projects which attempt to
parse Postfix logs are summarised (full details are available in
appendix~\refwithpage{other-parsers}), finishing with a review of
previously published research in this area.

@@

This paper and the program it describes are part of a larger project to
optimise a server's Postfix restrictions, generate statistics and graphs,
and provide a platform on which new restrictions can be trialled and
evaluated to see if they are beneficial in the fight against spam.  The
program parses Postfix logs and populates a database with the data gleaned
from those logs, providing a consistent and simple view of the logs which
future tools can utilise.  The gathered data can then be used to optimise
current anti-spam measures, provide a baseline to test new anti-spam
measures against, or to produce statistics showing how effective those
measures are.

@@

The database populated by this program provides a simple interface to
Postfix logs.  Although the interface is a database schema, it is in effect
quite similar to any other \API{} provided by shared code: it insulates
both user and provider of the \API{} from changes in the implementation of
the \API{}\@.  The algorithm implemented by the parser can be improved;
support can be added for earlier or later releases of Postfix; bugs can be
fixed or limitations removed from the parser; these changes will not cause
the user to be negatively impacted.  Statistics and/or graphs can be
generated from the database; new restrictions can be tested and the results
inspected; trends in the fight against spam can emerge from historical data
saved in the database; the parser remains the same as the usage adapts.
Using a database simplifies writing programs which need to interact with
the data in several ways:

@@

Storing the results in a database will also increase the efficiency of
using those results, as the logs need only be parsed once; indeed the
results may be used by someone with no access to the original logs.

@@

Postfix is a \MTA{} with the following design aims (in order of
importance): security, flexibility of configuration, scalability, and high
performance.  It features extensive, extensible, optional anti-spam
restrictions, allowing an administrator to deploy those restrictions which
they judge suitable for their site's needs, rather than a fixed set chosen
by Postfix's author.  These restrictions can be selectively applied,
combined and bypassed on a per-client, per-recipient or per-sender basis,
allowing varying levels of severity and/or permissiveness.  Postfix
leverages simple lookup tables to support arbitrarily complicated
user-defined sequences of restrictions and exceptions, with policy
servers\footnote{Policy servers will be explained in \sectionref{policy
servers}.} as the ultimate in flexibility.  Administrators can also supply
their own rejection messages to make it clear to senders why exactly their
mail was rejected.  Unfortunately this flexibility has a cost: complexity
in the logs generated.  While it is easy to use standard Unix text
processing utilities to determine the fate of an individual email,
following the journey an email takes through Postfix can be quite
difficult.  The logs tend to follow a 90\%-10\% pattern: 90\% of the time
the journey is simple, but the other 10\% of the time requires 90\% of the
code.\footnote{These numbers don't have a solid scientific basis, they're
based on gut feeling from writing and debugging the software.}

@@

    \item Greylisting~\cite{greylisting} is a technique that temporarily
        rejects mail when the triple of (sender address, recipient address,
        remote \IP{} address) is unknown; on second and subsequent delivery
        attempts from that triple the mail will be accepted.  The
        assumption is that maintaining a list of failed addresses and
        retrying after a temporary failure is uneconomical for a spam
        sender, but that a legitimate mail server must retry.  Sadly spam
        senders are using increasingly complex and well written programs to
        distribute spam, frequently using an \ISP{} provided \SMTP{} server
        from a compromised machine on the \ISP{}'s network.  Greylisting
        will slowly become less useful, but it does block a large
        percentage of spam mail at the moment; the most effective
        restrictions over the \numberOFlogFILES{} log files used in testing
        the parser are shown in table~\refwithpage{Summary of rejections}.
        Greylisting is obviously worth using, at least at the moment,
        particularly when you factor in Greylisting's position as the final
        restriction which a mail must overcome:\footnote{Greylisting is the
        final restriction a mail must overcome in the configuration used on
        the mail server the logs were obtained from; an administrator is
        free to use Greylisting at whichever position in the restriction
        list they feel is most appropriate for their mail system.}
        Greylisting only takes effect for mails which have passed every
        other restriction.

@@

    \item The logs are whole and complete: nothing has been removed, either
        deliberately or accidentally (e.g.\ log rotation gone awry, file
        system filling up, logging system unable to cope with the volume of
        logs).  On a well run system it is extremely unlikely that any of
        these problems will arise, though it is of course possible,
        particularly when suffering a deluge of spam or a mail loop.

@@

    \item The parser is developed using real world logs rather than the
        idealised logs someone would naturally envisage reading the source
        code.

@@

It should be clear from the earlier Postfix background (\sectionref{postfix
background}) that logs produced by Postfix are not fixed; they vary widely
from host to host, depending on the set of restrictions chosen by the
administrator.  With this in mind, one of the parser's design aims was to
make adding new rules as easy as possible, to enable administrators to
properly parse their log lines.  To enable this the parser is divided into
three parts:

@@

Parsing in \LMA{} requires major changes to the code to parse new log lines
or extract additional data.  It does not appear to deal with any of the
complications discussed in \sectionref{complications} (initial
complications) and \sectionref{additional complications} (additional
complications).  It does not differentiate between different types of
rejections, so it is not suitable for the purposes of this project; the
data about which restriction caused the rejection is discarded, whereas an
important goal of this project is to retain that data to aid optimisation
and evaluation of Postfix restrictions.  \LMA{} fails to parse Postfix log
files generated on Solaris hosts because the fields added automatically
differ from those added on Linux hosts; logs from Solaris hosts (and
possibly other operating systems) thus require preprocessing before parsing
by \LMA{}.  Once the preprocessing has been performed on the
\numberOFlogFILES{} test log files \LMA{} parses the log files without
complaint, though it does produce 32 entries in its output file for every
rejection in the input log file; it also missed some 40\% of delivered
mail.  Once these glaring deficiencies were discovered the author did not
spend more time checking the results.

>>>>

This section provides background information helpful in understanding the
remainder of the paper.  It begins with a discussion of the motivation
underlying the project, followed by some technical information: the use of
a database as an \API{}\@; a brief introduction to \SMTP{}\@; a longer
introduction to Postfix, concentrating on the topics most relevant to this
paper, namely Postfix anti-spam restrictions and policy servers.  The
assumptions made in designing and implementing the parser are explained, as
are the conventions used in this paper.  Other projects which attempt to
parse Postfix log files are summarised (full details are available in
appendix~\refwithpage{other-parsers}), finishing with a review of
previously published research in this area.

@@

This paper and the program it describes are part of a larger project to
optimise a server's Postfix restrictions, generate statistics and graphs,
and provide a platform on which new restrictions can be trialled and
evaluated to see if they are beneficial in the fight against spam.  The
program parses Postfix log files and populates a database with the data
gleaned from those log files, providing a consistent and simple view of the
log files which future tools can utilise.  The gathered data can then be
used to optimise current anti-spam measures, provide a baseline to test new
anti-spam measures against, or to produce statistics showing how effective
those measures are.

@@

The database populated by this program provides a simple interface to
Postfix log files.  Although the interface is a database schema, it is in
effect quite similar to any other \API{} provided by shared code: it
insulates both user and provider of the \API{} from changes in the
implementation of the \API{}\@.  The algorithm implemented by the parser
can be improved; support can be added for earlier or later releases of
Postfix; bugs can be fixed or limitations removed from the parser; these
changes will not cause the user to be negatively impacted.  Statistics
and/or graphs can be generated from the database; new restrictions can be
tested and the results inspected; trends in the fight against spam can
emerge from historical data saved in the database; the parser remains the
same as the usage adapts.  Using a database simplifies writing programs
which need to interact with the data in several ways:

@@

Storing the results in a database will also increase the efficiency of
using those results, as the log files need only be parsed once; indeed the
results may be used by someone with no access to the original log files.

@@

Postfix is a \MTA{} with the following design aims (in order of
importance): security, flexibility of configuration, scalability, and high
performance.  It features extensive, extensible, optional anti-spam
restrictions, allowing an administrator to deploy those restrictions which
they judge suitable for their site's needs, rather than a fixed set chosen
by Postfix's author.  These restrictions can be selectively applied,
combined and bypassed on a per-client, per-recipient or per-sender basis,
allowing varying levels of severity and/or permissiveness.  Postfix
leverages simple lookup tables to support arbitrarily complicated
user-defined sequences of restrictions and exceptions, with policy
servers\footnote{Policy servers will be explained in \sectionref{policy
servers}.} as the ultimate in flexibility.  Administrators can also supply
their own rejection messages to make it clear to senders why exactly their
mail was rejected.  Unfortunately this flexibility has a cost: complexity
in the log files generated.  While it is easy to use standard Unix text
processing utilities to determine the fate of an individual email,
following the journey an email takes through Postfix can be quite
difficult.  The log files tend to follow a 90\%-10\% pattern: 90\% of the
time the journey is simple, but the other 10\% of the time requires 90\% of
the code.\footnote{These numbers don't have a solid scientific basis,
they're based on gut feeling from writing and debugging the software.}

@@

    \item Greylisting~\cite{greylisting} is a technique that temporarily
        rejects mail when the triple of (sender address, recipient address,
        remote \IP{} address) is unknown; on second and subsequent delivery
        attempts from that triple the mail will be accepted.  The
        assumption is that maintaining a list of failed addresses and
        retrying after a temporary failure is uneconomical for a spam
        sender, but that a legitimate mail server must retry.  Sadly spam
        senders are using increasingly complex and well written programs to
        distribute spam, frequently using an \ISP{} provided \SMTP{} server
        from a compromised machine on the \ISP{}'s network.  Greylisting
        will slowly become less useful, but it does block a large
        percentage of spam mail at the moment; the most effective
        restrictions over the \numberOFlogFILES{} log files used in testing
        the parser are shown in table~\refwithpage{Summary of rejections}.
        Greylisting is obviously worth using, at least at the moment,
        particularly when you factor in Greylisting's position as the final
        restriction which a mail must overcome:\footnote{Greylisting is the
        final restriction a mail must overcome in the configuration used on
        the mail server the log files were obtained from; an administrator
        is free to use Greylisting at whichever position in the restriction
        list they feel is most appropriate for their mail system.}
        Greylisting only takes effect for mails which have passed every
        other restriction.

@@

    \item The log files are whole and complete: nothing has been removed,
        either deliberately or accidentally (e.g.\ log rotation gone awry,
        file system filling up, logging system unable to cope with the
        volume of log messages).  On a well run system it is extremely
        unlikely that any of these problems will arise, though it is of
        course possible, particularly when suffering a deluge of spam or a
        mail loop.

@@

    \item The parser is developed using real world log files rather than
        the idealised log files someone would naturally envisage reading
        the source code.

@@

It should be clear from the earlier Postfix background (\sectionref{postfix
background}) that log files produced by Postfix are not fixed; they vary
widely from host to host, depending on the set of restrictions chosen by
the administrator.  With this in mind, one of the parser's design aims was
to make adding new rules as easy as possible, to enable administrators to
properly parse their log lines.  To enable this the parser is divided into
three parts:

@@

Parsing in \LMA{} requires major changes to the code to parse new log lines
or extract additional data.  It does not appear to deal with any of the
complications discussed in \sectionref{complications} (initial
complications) and \sectionref{additional complications} (additional
complications).  It does not differentiate between different types of
rejections, so it is not suitable for the purposes of this project; the
data about which restriction caused the rejection is discarded, whereas an
important goal of this project is to retain that data to aid optimisation
and evaluation of Postfix restrictions.  \LMA{} fails to parse Postfix log
files generated on Solaris hosts because the fields added automatically
differ from those added on Linux hosts; log files from Solaris hosts (and
possibly other operating systems) thus require preprocessing before parsing
by \LMA{}.  Once the preprocessing has been performed on the
\numberOFlogFILES{} test log files \LMA{} parses the log files without
complaint, though it does produce 32 entries in its output file for every
rejection in the input log file; it also missed some 40\% of delivered
mail.  Once these glaring deficiencies were discovered the author did not
spend more time checking the results.

### 2008-05-20 doc/logparser-conclusion.tex
Moved \appendix to a more appropriate place. Added postfix bibliography entry; I should refer to it from logparser.pdf.
:%s/logs/log files/gc

Parsing Postfix logs appears at first sight to be an almost trivial task,
especially if one has previous experience in parsing logs, but it turns out
to be a much more taxing project than initially expected.  The variety and
breadth of log lines produced by Postfix is quite surprising, because a
quick survey of sample log files gives the impression that the number of
distinct log lines is quite small; this impression is due to the uneven
distribution exhibited by logs produced in normal operation (see
graph~\refwithpage{rule hits graph} for a vivid illustration of this).

@@

The real difficulties arise once the parser is successfully dealing with
90\% of the log lines, as the irregularities and complications previously
explained only begin to become apparent once the vast majority of log lines
have been parsed successfully.  Adding new rules to deal with numerous,
infrequently occurring log lines is a simple task, albeit tiresome (though
alleviated by the tool described in \sectionref{adding new actions}),
whereas dealing with mails where information appears to be missing is much
more grueling.  Trawling through the logs, looking for something out of the
ordinary, possibly hundreds or even thousands of lines away from the last
mention of the queueid in question, is extremely time consuming and error
prone.  In some cases the task is not to spot the line which is unusual,
but to spot that a line normally present is missing, i.e.\ to realise that
one line amongst hundreds is absent.  In all cases the evidence must be
used to construct a hypothesis to explain the irregularities, and that
hypothesis must then be tested in the parser; if successful, the parser
must be modified to deal with the irregularities, without adversely
affecting the existing functionality.  The complications described in
\sectionref{additional complications} were solved in the order they are
described in, and that order closely resembles the frequency with which
they occur; the most frequently occurring complications dominate the
warning messages produced, and so naturally they are the first
complications to be dealt with.

@@

\appendix

>>>>

Parsing Postfix log files appears at first sight to be an almost trivial
task, especially if one has previous experience in parsing log files, but
it turns out to be a much more taxing project than initially expected.  The
variety and breadth of log lines produced by Postfix is quite surprising,
because a quick survey of sample log files gives the impression that the
number of distinct log lines is quite small; this impression is due to the
uneven distribution exhibited by log lines produced in normal operation
(see graph~\refwithpage{rule hits graph} for a vivid illustration of this).

@@

The real difficulties arise once the parser is successfully dealing with
90\% of the log lines, as the irregularities and complications previously
explained only begin to become apparent once the vast majority of log lines
have been parsed successfully.  Adding new rules to deal with numerous,
infrequently occurring log lines is a simple task, albeit tiresome (though
alleviated by the tool described in \sectionref{adding new actions}),
whereas dealing with mails where information appears to be missing is much
more grueling.  Trawling through the log files, looking for something out
of the ordinary, possibly hundreds or even thousands of lines away from the
last mention of the queueid in question, is extremely time consuming and
error prone.  In some cases the task is not to spot the line which is
unusual, but to spot that a line normally present is missing, i.e.\ to
realise that one line amongst hundreds is absent.  In all cases the
evidence must be used to construct a hypothesis to explain the
irregularities, and that hypothesis must then be tested in the parser; if
successful, the parser must be modified to deal with the irregularities,
without adversely affecting the existing functionality.  The complications
described in \sectionref{additional complications} were solved in the order
they are described in, and that order closely resembles the frequency with
which they occur; the most frequently occurring complications dominate the
warning messages produced, and so naturally they are the first
complications to be dealt with.

@@

### 2008-05-20 doc/logparser-coverage.tex
:%s/logs/log files/gc

The state tables contain entries for mails not yet delivered when the
parser finishes execution.  Ideally all they should contain are mails which
are awaiting delivery after the period covered by the logs, though they may
also contain mails whose initial entries are not contained in the logs.
Any other entries are evidence of a failure in parsing or an aberration in
the logs.  After parsing the \numberOFlogFILES{} test log files the state
tables contain 18 entries, breaking down into:

@@

    \item 1 connection which started only seconds before the logs ended and
        had not yet been fully transferred from client to server.

    \item 1 mail which had been accepted only seconds before the logs ended
        and had not yet been delivered.

    \item 9 mails whose initial log entries were not present in the logs.

@@

Whereas verifying by inspection that the parser correctly deals with all
60,721,709 lines in the test logs is infeasible, verifying a subset of
those logs is a tractable, if extremely time consuming, task.  A sample of
logs was obtained by randomly selecting a log file:

@@

Parser coverage is divided into two topics in this section: log lines
covered and mails covered.  The former is initially more important, as the
parser must successfully parse every line if it is to be complete, but
subsequently the latter takes precedence because reproducing the path a
mail takes through Postfix is the aim of the parser.  Increasing the
percentage of log lines parsed is relatively simple and non-intrusive:
adding new rules or modifying existing rules is simplified by the
separation of rules, actions and framework.  Improving the logical coverage
is harder, as the actions taken by Postfix must be reconstructed by the
author, and the new sequence of actions integrated into the existing model
without breaking the existing parsing.  Detecting a deficiency in the
parsing algorithm is also significantly harder than detecting unparsed log
lines, as the parser will warn about any unparsed line, whereas discovering
a flaw in the parser requires understanding of the warnings produced and
the mails remaining in the state table.  Rectifying a flaw in the parser
requires an understanding of both the parser and Postfix's logs, and
investigative work to determine the cause of the deficiency, followed by
further examination of the logs in developing a solution.

>>>>

The state tables contain entries for mails not yet delivered when the
parser finishes execution.  Ideally all they should contain are mails which
are awaiting delivery after the period covered by the log files, though
they may also contain mails whose initial entries are not contained in the
log files.  Any other entries are evidence of a failure in parsing or an
aberration in the log files.  After parsing the \numberOFlogFILES{} test
log files the state tables contain 18 entries, breaking down into:

@@

    \item 1 connection which started only seconds before the log files
        ended and had not yet been fully transferred from client to server.

    \item 1 mail which had been accepted only seconds before the log files
        ended and had not yet been delivered.

    \item 9 mails whose initial log entries were not present in the log
        files.

@@

Whereas verifying by inspection that the parser correctly deals with all
60,721,709 lines in the test log files is infeasible, verifying a subset of
those log files is a tractable, if extremely time consuming, task.  A
sample of log lines was obtained by randomly selecting a log file:

@@

Parser coverage is divided into two topics in this section: log lines
covered and mails covered.  The former is initially more important, as the
parser must successfully parse every line if it is to be complete, but
subsequently the latter takes precedence because reproducing the path a
mail takes through Postfix is the aim of the parser.  Increasing the
percentage of log lines parsed is relatively simple and non-intrusive:
adding new rules or modifying existing rules is simplified by the
separation of rules, actions and framework.  Improving the logical coverage
is harder, as the actions taken by Postfix must be reconstructed by the
author, and the new sequence of actions integrated into the existing model
without breaking the existing parsing.  Detecting a deficiency in the
parsing algorithm is also significantly harder than detecting unparsed log
lines, as the parser will warn about any unparsed line, whereas discovering
a flaw in the parser requires understanding of the warnings produced and
the mails remaining in the state table.  Rectifying a flaw in the parser
requires an understanding of both the parser and Postfix's log files, and
investigative work to determine the cause of the deficiency, followed by
further examination of the log files in developing a solution.

### 2008-05-20 doc/logparser-database-schema.tex
Added cluster_groups to the rules table description and example rule. Improved Splunk review. Improved summary table in review of other parsers conclusion. Minor improvements in parser reviews. Improved review of other parsers conclusion.
:%s/logs/log files/gc

The data saved from parsing the logs is also divided into two tables as
described below: connections and results.  The connections table contains a
row for every mail accepted and every connection where there was a
rejection; the individual fields will be described in
\sectionref{connections table}.  The results table has one or more rows per
row in the connections table; the fields will be covered in detail in
\sectionref{results table}.

@@

            \item [POSTFIX\_RELOAD] The administrator has instructed
                Postfix to start or stop, and all existing \daemon{smtpd}
                processes will be terminated.  This does not negatively
                impact on the logs or mail queued by Postfix for delivery.

@@

@@

The table containing the rules used by the parser and both tables
containing the data extracted from the Postfix logs were described, with
the purpose of each field discussed in detail.  A clear, comprehensible
schema is essential when using the extracted data; it's more important when
using the data than when storing it, because storing the data is a
write-once operation, whereas utilising the data requires frequent
searching, sorting and manipulation of the data to produce customised
reports and/or statistics.

>>>>

The data saved from parsing the log files is also divided into two tables
as described below: connections and results.  The connections table
contains a row for every mail accepted and every connection where there was
a rejection; the individual fields will be described in
\sectionref{connections table}.  The results table has one or more rows per
row in the connections table; the fields will be covered in detail in
\sectionref{results table}.

@@

            \item [POSTFIX\_RELOAD] The administrator has instructed
                Postfix to start or stop, and all existing \daemon{smtpd}
                processes will be terminated.  This does not negatively
                impact on the log files or mail queued by Postfix for
                delivery.

@@

    \item [cluster\_group] A reference to the \texttt{cluster\_group}
        table.  That table is used by the Decision Tree algorithm described
        in a separate document.

@@

The table containing the rules used by the parser and both tables
containing the data extracted from the Postfix log files were described,
with the purpose of each field discussed in detail.  A clear,
comprehensible schema is essential when using the extracted data; it's more
important when using the data than when storing it, because storing the
data is a write-once operation, whereas utilising the data requires
frequent searching, sorting and manipulation of the data to produce
customised reports and/or statistics.

### 2008-05-20 doc/logparser-graphs.tex
:%s/logs/log files/gc

The sequence described above occurs extremely rapidly because Postfix does
not have to deliver the mail to an external system, so mails are delivered,
bounced and generated as fast as the disks can keep up, resulting in a huge
volume of logs.

@@

The difference between logs with a mail loop and logs without can be seen
in table~\refwithpage{Execution time increase for different rule orderings}
showing the increases for the different rule orderings and combinations of
log files:

>>>>

The sequence described above occurs extremely rapidly because Postfix does
not have to deliver the mail to an external system, so mails are delivered,
bounced and generated as fast as the disks can keep up, resulting in a huge
volume of log lines.

@@

The difference between log files with a mail loop and log files without can
be seen in table~\refwithpage{Execution time increase for different rule
orderings} showing the increases for the different rule orderings and
combinations of log files:

### 2008-05-20 doc/logparser-introduction.tex
:%s/logs/log files/gc

Most mail server administrators will have performed some basic processing
of the logs produced by their mail server at one time or another, whether
it was to debug a problem, explain to a user why their mail is being
rejected, or check whether new anti-spam measures are working.  The more
adventurous will have generated statistics to show how many hits each of
their anti-spam measures has gotten in the last week, and possibly even
generated some graphs to clearly illustrate the point to management or
users.\footnote{This was the author's first real foray into processing
Postfix logs.}  Very few will have performed in-depth parsing and analysis
of their logs, where the parsing must correlate the log lines
per-connection or per-queueid rather than processing log lines
independently.  One of the barriers to this type of processing is the
unstructured nature of Postfix logs, where each log line was added on an ad
hoc basis as a requirement was discovered or new functionality was
added.\footnote{A history of all changes made to Postfix is distributed
with the source code, available from \url{http://www.postfix.org/}} Further
complication arises because the set of rejection messages is not fixed: new
messages can be added by the administrator with custom checks; every
\RBL{}\footnote{This document is supplied with a glossary, see
appendix~\ref{Glossary}.} returns a different explanatory message; policy
servers may log different messages depending on the characteristics of the
connection; there are many ways in which the log lines may differ between
servers, even within the same organisation --- servers may be configured
differently, or running different version of Postfix.  This paper documents
the difficult process of parsing Postfix logs, presenting \PLP{}, a program
which parses logs and places the resulting data into a database for later
use.  The gathered data can then be used to optimise current anti-spam
measures, provide a baseline to test new anti-spam measures against, or to
produce statistics showing how effective those measures are.  There are
numerous other uses for such data: improving server performance by
identifying troublesome destinations and reconfiguring appropriately;
identifying regular high volume uses (e.g.\ customer newsletters) and
restricting those uses to off-peak times; detecting virus outbreaks which
propagate via email; as a base for billing customers on a shared server.
Preserving the raw data enables users to develop a multitude of uses far
beyond those conceived of by the author.

@@

This algorithm requires a database for storing both the rules used when
parsing and the results gleaned from parsing.  The database schema used is
described in section~\refwithpage{database schema}, explaining in detail
the tables used for storing the data gleaned from the logs and the table
that stores the rules.

>>>>

Most mail server administrators will have performed some basic processing
of the log files produced by their mail server at one time or another,
whether it was to debug a problem, explain to a user why their mail is
being rejected, or check whether new anti-spam measures are working.  The
more adventurous will have generated statistics to show how many hits each
of their anti-spam measures has gotten in the last week, and possibly even
generated some graphs to clearly illustrate the point to management or
users.\footnote{This was the author's first real foray into processing
Postfix log files.}  Very few will have performed in-depth parsing and
analysis of their log files, where the parsing must correlate the log lines
per-connection or per-queueid rather than processing log lines
independently.  One of the barriers to this type of processing is the
unstructured nature of Postfix log files, where each log line was added on
an ad hoc basis as a requirement was discovered or new functionality was
added.\footnote{A history of all changes made to Postfix is distributed
with the source code, available from \url{http://www.postfix.org/}} Further
complication arises because the set of rejection messages is not fixed: new
messages can be added by the administrator with custom checks; every
\RBL{}\footnote{This document is supplied with a glossary, see
appendix~\ref{Glossary}.} returns a different explanatory message; policy
servers may log different messages depending on the characteristics of the
connection; there are many ways in which the log lines may differ between
servers, even within the same organisation --- servers may be configured
differently, or running different version of Postfix.  This paper documents
the difficult process of parsing Postfix log files, presenting \PLP{}, a
program which parses log files and places the resulting data into a
database for later use.  The gathered data can then be used to optimise
current anti-spam measures, provide a baseline to test new anti-spam
measures against, or to produce statistics showing how effective those
measures are.  There are numerous other uses for such data: improving
server performance by identifying troublesome destinations and
reconfiguring appropriately; identifying regular high volume uses (e.g.\
customer newsletters) and restricting those uses to off-peak times;
detecting virus outbreaks which propagate via email; as a base for billing
customers on a shared server.  Preserving the raw data enables users to
develop a multitude of uses far beyond those conceived of by the author.

@@

This algorithm requires a database for storing both the rules used when
parsing and the results gleaned from parsing.  The database schema used is
described in section~\refwithpage{database schema}, explaining in detail
the tables used for storing the data gleaned from the log files and the
table that stores the rules.

### 2008-05-20 doc/logparser-limitations.tex
:%s/logs/log files/gc

    \item The parser does not distinguish between logs produced by
        different sources when parsing; all results will be saved to the
        same database.  This may be viewed as an advantage, as logs from
        different sources can be combined in the same database, or it may
        be viewed as a limitation as there is no facility to distinguish
        between logs from different sources in the same database.  If the
        results of parsing logs from different sources must remain
        separate, the parser can easily be instructed to use a different
        database to store the results in.

>>>>

    \item The parser does not distinguish between log files produced by
        different sources when parsing; all results will be saved to the
        same database.  This may be viewed as an advantage, as log files
        from different sources can be combined in the same database, or it
        may be viewed as a limitation as there is no facility to
        distinguish between log files from different sources in the same
        database.  If the results of parsing log files from different
        sources must remain separate, the parser can easily be instructed
        to use a different database to store the results in.
